material and methods

### 2018-01-29 man/ms_1_material_methods.Rmd
update text

This species reaches its southernmost European limit at Sierra Nevada, a high-mountain range located in southern Spain (37°N, 3°W) with elevations of between 860 m and 3482 m *a.s.l.*. The climate is Mediterranean, characterized by cold winters and hot summers, with pronounced summer drought (July-August). There are eight oak patches (2400 Has) identified (**Figure 1**) in this mountain range, ranging between 1100 and 2000 m *a.s.l.* and generally associated to major river valleys. Sierra Nevada is considered a glacial refugia for deciduous *Quercus* species during glaciation [@Brewer2002; @Olalde2002; RodriguezSanchez2010] and these populations are considered as a rear edge of the habitat distribution, which is important in determining habitat responses to expected climate change [@Hampe2005].

The populations of Pyrenean oak forests at Sierra Nevada are considered relict forests [@Melendo2000, @Vivero2000], undergoing intensive anthropic use in the last few decades [@CamachoOlmedo2002; @Valbuena2010; @Valbuena2017]. In fact, the status of conservation of this species for southern Spain is "Vulnerable" [@Vivero2000]. The relict presence of this species in Sierra Nevada is related both to its genetic resilience as well as to its high intraspecific genetic diversity [@Valbuena2010]. However, they are also expected to suffer the impact of climate change, due to their climate requirements (wet summers). Thus, simulations of the climate change effects on this habitat forecast a reduction in suitable habitats for Sierra Nevada [@Benito2011].

@@

* :red_circle: Meter aquí algunos datos de sequia, similar a lo planteado por Gazol

:red_circle: `duda aqui` Varias referencias hablan de los años 2005 y 2012 como  extremadamente secos. Pero habría que hacer alguna referencia y/o análisis. Tengo dudas de si hemos de analizar (e incluir) que efectivamente los años 2005 y 2012 fueron caracterizados por un extrema sequía, por lo que habría que  incluyendo referencia a apéndice

* O quizá un apartado llamado Drought episodes (similar a esto https://www.nature.com/articles/srep28269)

@@

>>>>

This species reaches its southernmost European limit at Sierra Nevada, a high-mountain range located in southern Spain (37°N, 3°W) with elevations of between 860 m and 3482 m *a.s.l.*. The climate is Mediterranean, characterized by cold winters and hot summers, with pronounced summer drought (July-August). There are eight oak patches (2400 Has) identified (Figure 1) in this mountain range, ranging between 1100 and 2000 m *a.s.l.* and generally associated to major river valleys. Sierra Nevada is considered a glacial refugia for deciduous *Quercus* species during glaciation [@Brewer2002; @Olalde2002; @RodriguezSanchez2010] and these populations are considered as a rear edge of the habitat distribution, which is important in determining habitat responses to expected climate change [@Hampe2005].

The populations of Pyrenean oak forests at Sierra Nevada are considered relict forests [@Melendo2000; @Vivero2000], undergoing intensive anthropic use in the last few decades [@CamachoOlmedo2002; @Valbuena2010; @Valbuena2017]. In fact, the status of conservation of this species for southern Spain is "Vulnerable" [@Vivero2000]. The relict presence of this species in Sierra Nevada is related both to its genetic resilience as well as to its high intraspecific genetic diversity [@Valbuena2010]. However, they are also expected to suffer the impact of climate change, due to their climate requirements (wet summers). Thus, simulations of the climate change effects on this habitat forecast a reduction in suitable habitats for Sierra Nevada [@Benito2011].

@@

The years 2005 and 2012 were chosen as drought years for this study because they have been documented as two of the worst drought years particularly for the southern of the Iberian Peninsula [@GarciaHerrera2007; @Gouveia2015; @Trigo2013] and the were charecterized by extreme drogught (Appendix S1). Aunque sabemos que existen otros años de sequía extrema (e.g. 1995, 1981) [@VicenteSerrano2014], no se han elegido por una limitación metodológica, esto es, nuestro objetivo es analizar el efecto de la sequía y la resiliencia a los eventos de sequía utilizando información de satélite y de dendrocronología, y la información de satélite utilizada (ver mas adelante) solamente está disponible desde el año 2000 en adelante.

```{r echo=FALSE, results='asis', eval=show_text}
cat("
* Ver ejemplo Gazol et al 2017 Plan Ecology
* un apartado llamado Drought episodes (similar a esto https://www.nature.com/articles/srep28269)
")
```

@@

```{r echo=FALSE, results='asis', eval=show_text}
cat("
SPEI values between 1 and −1 are considered normal, whereas values <−1 indicate drought and values <−1.5 indicate severe drought47. We defined a drought as a period where SPEI12 decreased to ≤−1.5.
")
```
